### 2022-04-18 introduction/introduction.tex
more

Integrating proof automation directly into the proof assistant is not an easy task.
Most proof assistants are based on variants of higher-order logic, which is more
expressive than first-order logic, but harder to automate as in general it does not even allow complete
proof calculi. To provide some level of automation, proof assistant developers
use the efficient first-order provers as follows. First, the proof goal is
translated from higher-order to first-order logic. Then the first-order prover
is run on the translated problem. If the first-order prover finds the proof, the
proof is reconstructed in the proof assistant. Modern assistants such as Coq \cite{bc-04-coq}, HOL Light \cite{jh-09-hol-light}
and Isabelle \cite{lc-88-isabelle} implement the approach using the \emph{hammers} CoqHammer
\cite{ck-18-coqhammer}, HOL(y)Hammer \cite{ku-15-holyhammer} and Sledgehammer
\cite{bn-10-sh}, respectively.

>>>>

Integrating proof automation directly into the proof assistant is not an easy task.
Most proof assistants are based on variants of higher-order logic, which is more
expressive than first-order logic, but harder to automate as in general it does not even allow complete
proof calculi. To provide some level of automation, proof assistant developers
use the efficient first-order provers as follows. First, the proof goal is
translated from higher-order to first-order logic. Then the first-order prover
is run on the translated problem. If the first-order prover finds the proof, the
proof is reconstructed in the proof assistant. Modern assistants such as Coq \cite{bc-04-coq}, HOL Light \cite{jh-09-hol-light},
and Isabelle \cite{lc-88-isabelle} implement the approach using the \emph{hammers} CoqHammer
\cite{ck-18-coqhammer}, HOL(y)Hammer \cite{ku-15-holyhammer}, and Sledgehammer
\cite{bn-10-sh}, respectively.

### 2022-04-18 pre/pre.tex
more

This calculus works on problems in clausified form and proves the problem by
refuting its negation. In practice this means that, to apply superposition to a
given problem, we first must negate its conjecture and clausify the whole
problem using some of the clausification algorithms
\cite{nw-01-small-cnf,rsv-16-vcnf}. The calculus then applies its inference
rules to the clauses, adding results of the inferences to the working set of
clauses. As this process is quite prolific, it also uses an order on terms which gives an indication of a
term's ``simplicity'' to determine
if some clauses can be simplified or eliminated from the working set.
Superposition is refutationally complete: This means that, if results of all
inferences are computed in a systematic, fair manner, the empty clause ($\bot$) will
eventually be derived for a provable problem. As first-order logic is
undecidable, no guarantees are given for unprovable problems.
Before introducing the rules of the calculus, we provide definitions of these and other
background concepts.

@@

Calculi from the resolution family, including superposition, perform inferences
only on unifiable terms. We say terms $s$ and $t$ are {\em unifiable} if there is a
substitution $\sigma$ such that $\sigmaterm{s} = \sigmaterm{t}$; we further say
$\sigma$ is a unifier. A \emph{unification constraint} $s \unif t$ is a pair of two terms of the same type. 
We always specify if the constraint should be interpreted as ordered or unordered.
We say a substitution $\sigma$ is \emph{more general} than $\varrho$ if there exists a substitution $\theta$
such that $\varrho = \theta\sigma$.
A \emph{most general unifier} ({MGU}) is a
unifier $\sigma$, such that for any other unifier $\theta$, there is a
substitution $\varrho$ such that $\theta = \varrho\sigma$. In first-order logic,
the MGU is unique up to variable renaming. For example, $\cst{f}(X) \unif \cst{f}(Y)$ yields the MGU 
$\{ X \mapsto Y\}$ or  $\{ Y \mapsto X \}$. First-order logic also admits efficient algorithms for computing MGUs \cite{hv-09-unifalgs}.

@@

The rules described above are used to generate new clauses from already derived
ones. Even though orders and selection are used to reduce the search space,
it still grows rather fast. To identify unnecessary clauses in the search
space, superposition features a \emph{redundancy criterion}
\cite[Sect.~4.2.2]{bg-01-resolution}. We say that a ground clause $C$ is
{\em redundant} in a ground set of clauses $N$ if it is entailed by a subset of
clauses in $N$ such that each clause in this subset is $\prec$-smaller than $C$. More
generally, a clause (ground or not) is redundant in set $N$ if every ground
instance of $C$ is redundant for the grounding of $N$. Removing redundant
clauses does not affect completeness of superposition.

Next to the four generating rules, superposition provers also use
simplification rules, which are justified by the redundancy criterion. These
rules effectively replace the pre\-mises with conclusions, and we denote them
using double bars. In theory, the rules work by adding conclusions and then
using them to show that premises are redundant, and can thus be eliminated. As
an example of such a rule, let us introduce rewriting (demodulation) of negative
literals (\infname{RN}):

@@

We say that a clause $C$ subsumes clause $D$ if there is a substitution $\sigma$
such that $\sigmacl{C} \subseteq D$. Subsumption is one of the most important
operations within a theorem prover. This rule does not adhere to the redundancy
criterion, but can be justified using other mechanisms \cite{wtrb-20-sat-framework}.

>>>>

This calculus works on problems in clausified form and proves the problem by
refuting its negation. In practice this means that, to apply superposition to a
given problem, we first must negate its conjecture and clausify the whole
problem using some of the clausification algorithms
\cite{nw-01-small-cnf,rsv-16-vcnf}. The calculus then applies its inference
rules to the clauses, adding results of the inferences to the working set of
clauses. As this process is quite prolific, it also uses an order on terms which gives an indication of a
term's ``simplicity'' to determine
if some clauses can be simplified or eliminated from the working set.
Superposition is refutationally complete: This means that, if results of all
inferences are computed in a systematic, fair manner, the empty clause ($\bot$) will
eventually be derived for a provable problem. As first-order logic is
undecidable, no guarantees are given for unprovable problems.
Before introducing the rules of the calculus, we provide definitions of the
background concepts.

@@

Calculi from the resolution family, including superposition, perform inferences
only on unifiable terms. We say terms $s$ and $t$ are {\em unifiable} if there is a
substitution $\sigma$ such that $\sigmaterm{s} = \sigmaterm{t}$; we further say
$\sigma$ is a unifier. A \emph{unification constraint} $s \unif t$ is a pair of two terms of the same type. 
We always specify if the constraint should be interpreted as ordered or unordered.
We say a substitution $\sigma$ is \emph{more general} than $\varrho$ if there exists a substitution $\theta$
such that $\varrho = \theta\sigma$.
A \emph{most general unifier} ({MGU}) is a
unifier $\sigma$, such that for any other unifier $\varrho$, $\sigma$ is more general than $\varrho$. In first-order logic,
the MGU is unique up to variable renaming. For example, $\cst{f}(X) \unif \cst{f}(Y)$ yields the MGU 
$\{ X \mapsto Y\}$ or  $\{ Y \mapsto X \}$. First-order logic also admits efficient algorithms for computing MGUs \cite{hv-09-unifalgs}.

@@

The rules described above are used to generate new clauses from already derived
ones. Even though orders and selection are used to reduce the search space,
it still grows rather fast. To identify unnecessary clauses in the search
space, superposition features a \emph{redundancy criterion}
\cite[Sect.~4.2.2]{bg-01-resolution}. We say that a ground clause $C$ is
{\em redundant} in a ground set of clauses $N$ if it is entailed by a subset of
clauses in $N$ such that each clause in this subset is $\succ$-smaller than $C$. More
generally, a clause (ground or not) is redundant in set $N$ if every ground
instance of $C$ is redundant for the grounding of $N$. Removing redundant
clauses does not affect completeness of superposition.

Next to the four generating rules, superposition provers also use
simplification rules, which are justified by the redundancy criterion. These
rules effectively replace the pre\-mises with conclusions, and we denote them
using double bars. In theory, the rules work by adding conclusions and then
using them to show that premises are redundant; after this, premises can be eliminated. As
an example of such a rule, let us introduce rewriting (demodulation) of negative
literals (\infname{RN}):

@@

We say that a clause $C$ subsumes clause $D$ if there is a substitution $\sigma$
such that $\sigmacl{C} \subseteq D$. Subsumption is one of the most important
operations of a theorem prover. This rule does not adhere to the redundancy
criterion, but can be justified using other mechanisms \cite{wtrb-20-sat-framework}.
